Bayesian Sparsification of Deep $\cplx$-valued networks

### 2020-02-07 paper/paper.tex
found proof for adding euler const, corrected proba estimate, reworded the netwroks description in mnist-like

We propse settng $\beta = 2$ for $\cplx$ variational dropout, to make the divergence
term depend only on $\alpha$.
% For $\beta = 2$ we add $\gamma$ to make the right-hand side nonegative. (proof missing)

@@

\subsection{Stagewise fitting and compression} % (fold)
\label{sub:fitting_and_compression}

@@

% on compresssion
Between \textit{``sparsify''} and \textit{``fine-tune''} stages we compute masks of non-zero
weights in each layer based on the relevance scores $\alpha$ (sec.~\ref{sub:dropout}). Since
the variational approximation factorizes into univariate distributions, whether a $\cplx$
or $\real$ parameter is zero or not is determined from individual relevance scores $
  \alpha = \tfrac{\sigma^2}{\lvert\mu\rvert^2}
$: non-zero if $\log \alpha \leq \tau$, zero otherwise. Figure~\ref{fig:hist__and__threshold__tradeoff}
displays the general shape of the distribution of $\log \alpha$ on the Fashion-MNIST dataset
in sec.~\ref{sub:mnist_like_datasets}. This shape appears to be typical in every experiment
conducted in this study: there always is a single mode within the moderately negative
range, and one or two modes to the right of zero. The relative mass of each mode changes
with $C$ and the higher the coefficient is, the more mass is shifted to the right. In every
experiment the trough between the peaks of the negative mode and the first positive one
is around zero.
%
The bottom plot on figure~\ref{fig:hist__and__threshold__tradeoff} is constructed without
running the ``fine-tune'' stage for $C=\tfrac1{2000}$, $\tfrac1{20}$, and $\frac1{200}$.
The latter setting yields one of the better performing models on the validation split in
sec.~\ref{sub:musicnet}. We can observe that by choosing $\tau > 0$ we end up in the
saturation region of the performance and gain only marginal improvement, while losing
in compression. Relative position compression curves shows that $C$ has much more substantial
impact on the target sparsity and performance, than the choice of $\tau$.

@@

% on chosing the threshold
The threshold $\tau$ is picked so, that the remaining \textit{``non-zero''} parameters
are within $\delta$ relative tolerance of their mean with high probability under the
approximate posterior $q_\theta(W)$. Since \eqref{eq:c-gauss-vi} is Gaussian, $
  \tfrac{\lvert w - \mu \rvert^2}
        {\alpha \lvert \mu \rvert^2}
$ is $\chi^2_k$ distributed with shape $k=1$ ($\real$) or $2$ ($\cplx$), whence
\begin{equation} \label{eq:prob_relevant}
  \Pr\bigl(
    \lvert w - \mu \rvert \leq \delta \lvert \mu \rvert
  \bigr)
    = \Pr\biggl(
      \frac{\lvert w - \mu \rvert^2}{\alpha \lvert \mu \rvert^2}
        \leq \frac{\delta^2}{\alpha}
    \biggr)
    = \chi^2_k\biggl(\frac{\delta^2}{\alpha}\biggr)
  \,.
\end{equation}
For a reasonably generous $\delta = 25\%$ tolerance all values of $\log \alpha$ below
$-2.5$ yield a relatively high chance of a non-zero parameter. Setting $\tau = 3$
\citep{molchanov_variational_2017,kingma_variational_2015} is, in our opinion, detrimental
since it retains parameters $W \sim q_\theta$, standard deviation of which significantly
exceeds the magnitude of $\mu$, which is typically used as a stand-in for $W$ when the
network is run in inference mode. Given that $q_\theta$ is merely an approximation,
we pick $\tau = -\tfrac12$ as the sparsification threshold for $\log \alpha$ in our
experiments.

% subsection fitting_and_compression (end)

\subsection{Metrics and performance} % (fold)
\label{sub:metrics_and_performance}

% perfromance metrics
We use accuracy as the performance metric for multi-class image classification on MNIST-like
and CIFAR10 datasets. Similarly to \citet{trabelsi_deep_2017} the performance in MusicNet
experiments is measured with (pooled) average precision: predicted scores and target binary
labels for each output are pooled, and average precision is computed based on the precision-recall
curve as if the task was single-output. The evaluation is done after ``dense'' stage, before
the beginning and upon termination of the ``fine-tune'' stage.

@@

% subsection metrics_and_performance (end)

@@

% model architectures
% discussion of TwoLayerDenseModel (TLDM) and SimpleConvModel (SCM)
 the following :
To compactly describe the $\real$/$\cplx$ networks we study in this experiment we use
the following notation: $c(n, k, s)$ is a convolution with $n$ output channels, kernel
width $k$ and stride $s$, and $d(n)$ is a dense layer with $n$ outputs. The values of
$n_\cdot$ depend on the input features and are summarized in table~\ref{tbl:mnist-like-model-spec}.
\begin{itemize}
  \item \textit{SimpleConvModel} (SCM): $2d$ convnet with Average pooling ($2\times 2$)
  with layers $c(n_1, 5, 1) \to c(n_2, 5, 1) \to d(n_3) \to d(n_\mathtt{out})$
  \item \textit{TwoLayerDenseModel} (TLDM): wide shallow network with structure
  $d(n_1) \to d(n_\mathtt{out})$
\end{itemize}

\begin{table}[hb]
  \centering
  \caption{Architecture parameters for models studied on MNIST-like datasets.}
  \label{tbl:mnist-like-model-spec}
  \begin{tabular}{lrr}
    \toprule
    $(n_1, n_2, n_3)$  & $\real$-\textit{SCM} & $\cplx$-\textit{SCM} \\
    \midrule
    (\texttt{raw}) & $20,  50,  500$ & $10, 25, 250$ \\
    (\texttt{fft}) & $40, 100, 1000$ & $20, 50, 500$ \\
    \bottomrule
  \end{tabular}
  \begin{tabular}{lrr}
    \toprule
    $(n_1)$  & $\real$-\textit{TLDM} & $\cplx$-\textit{TLDM} \\
    \midrule
    (\texttt{raw}) & $4096$ & $2048$ \\
    (\texttt{fft}) & $8192$ & $4096$ \\
    \bottomrule
  \end{tabular}
\end{table}

@@

We use training from sec.~\ref{sub:fitting_and_compression} with epochs lasting for $12$,
$32$ and $50$ epochs, respectively. Early stopping is used only during the ``fine-tune''
stage. Preliminary experiments with the model without sparsification have shown that early
stopping almost always terminated the training process within the first $10-20$ epochs,
and that the validation performance peaked at about $10-15$ epochs, much sooner than $200$
epoch, used by \citet{trabelsi_deep_2017}. Besides stagewise training and shorter ``dense''
stage, other important deviations from their setup include:
\begin{itemize}
  \item shifting lower frequencies to the centre of the window to maintain locality after
  applying FFT (much like in sec.~\ref{sub:mnist_like_datasets})
  \item clipping $\ell_2$ norm of parameters' gradients to $0.05$
  \item faster learning rate reduction schedule (sec.~\ref{sub:fitting_and_compression})
\end{itemize}
Each full three-stage experiment is replicated five times to take random effects into account,
e.g. random network initialization prior to ``dense'' stage, randomization of batches within
each epoch during SGD, and stochasticity of Bayesian forward pass at ``sparsify'' stage.

>>>>

We set $\beta = 2$ to make the divergence term depend only on $\alpha$ and add $\gamma$
so that the right-hand side is nonegative \citep[eq.(84)]{lapidoth_capacity_2003}.

@@

\subsection{Stagewise fitting and sparsification} % (fold)
\label{sub:fitting_and_sparsification}

@@

% on compresssion
Between \textit{``sparsify''} and \textit{``fine-tune''} stages we compute masks of non-zero
weights in each layer based on the relevance scores $\alpha$ (sec.~\ref{sub:dropout}). Since
the $q_\theta$ factorizes into univariate distributions, whether a $\cplx$ or $\real$ parameter
is zero or not is determined from $\alpha = \tfrac{\sigma^2}{\lvert\mu\rvert^2}$: non-zero
if $\log \alpha \leq \tau$, zero otherwise. Figure~\ref{fig:hist__and__threshold__tradeoff}
(top) displays the general shape of the distribution of $\log \alpha$ on the Fashion-MNIST
dataset in sec.~\ref{sub:mnist_like_datasets}. This shape appears to be typical in every
experiment conducted in this study: there always is a single mode within the moderately
negative range, and one or two modes to the right of zero. In every experiment the trough
between the peaks of the negative mode and the first positive one is around zero. The
relative mass of each mode changes with $C$ and the higher the coefficient is, the more
mass is shifted to the right.
%
The bottom graph on figure~\ref{fig:hist__and__threshold__tradeoff} is constructed for
$C \in \{\tfrac1{20}, \frac1{200}\}$ before running the ``fine-tune'' stage. From the
relative positions of the compression curves on this plot and from the results of other
experiments we conclude that $C$ has much more substantial impact on the sparsity and
performance, than the choice of $\tau$. Furthermore, for $\tau > 0$ the performance quickly
saturates and while compression continues to drop.

@@

% on chosing the threshold
The threshold $\tau$ is picked so, that the remaining \textit{``non-zero''} parameters
are within $\delta$ relative tolerance of their mean with high probability under the
approximate posterior $q_\theta(W)$. Since $W$ is $\real$- or a proper $\cplx$-Gaussian,
the random variable $
  \tfrac{k \lvert W - \mu \rvert^2}
        {\alpha \lvert \mu \rvert^2}
$ is $\chi^2_k$ distributed with $k=1$ ($\real$) or $2$ ($\cplx$), whence
\begin{equation} \label{eq:prob_relevant}
  \Pr\bigl(
    \lvert w - \mu \rvert \leq \delta \lvert \mu \rvert
  \bigr)
    = \Pr\biggl(
      \frac{k \lvert w - \mu \rvert^2}{\alpha \lvert \mu \rvert^2}
        \leq \frac{\delta^2 k}{\alpha}
    \biggr)
    = \chi^2_k\biggl(\frac{\delta^2 k}{\alpha}\biggr)
  \,.
\end{equation}
For a generous $\delta = 50\%$ tolerance all values of $\log \alpha$ below $-2.5$ yield
at least $90\%$ chance of a non-zero $\real$/$\cplx$ parameter. In our opinion, $\tau = 3$
\citep{molchanov_variational_2017,kingma_variational_2015} retains parameters with randomness
that significantly exceeds the magnitude of the mean, although $\mu$ is typically used as
a stand-in for $W$ when the network runs in inference mode. Given that $q_\theta$ is merely
an approximation, we pick $\tau = -\tfrac12$ to exclude noisy parameters and encourage
higher sparsity.

% subsection fitting_and_sparsification (end)

\subsection{Metrics and compression} % (fold)
\label{sub:metrics_and_compression}

% perfromance metrics
We use accuracy as the performance metric for multi-class image classification on MNIST-like
and CIFAR10 datasets. Similarly to \citet{trabelsi_deep_2017} performance in experiments with
MusicNet is measured with (pooled) average precision: predicted scores and target binary labels
for each output are pooled, and average precision is computed based on the precision-recall
curve as if the task was single-output. The evaluation is done after ``dense'' stage, before
the beginning and upon termination of the ``fine-tune'' stage.

@@

% subsection metrics_and_compression (end)

@@

% model architectures
% discussion of TwoLayerDenseModel (TLDM) and SimpleConvModel (SCM)
The $\real$/$\cplx$ networks, studied in this experiment, have been chosen for the purpose 
of illustrating the compression and understanding the effects of experiment parameters.
\textit{SimpleConvModel} (SCM) is $2d$ ReLU convolutional networks with average pooling
($2\times 2$) after non-linearities with two convolutions (kernel $5\times 5$, stride $1$)
with filters $n_1 \to n_2$ followed by dense layers $n_3 \to n_\mathtt{out}$. The second
network, \textit{TwoLayerDenseModel} (TLDM), is a wide dense network $n_1 \to n_\mathtt{out}$.
The values of $n_\cdot$ depend on the input features and are summarized in table~\ref{tbl:mnist-like-model-spec}.

\begin{table}[hb]
  \centering
  \caption{Architecture parameters for models studied on MNIST-like datasets.}
  \label{tbl:mnist-like-model-spec}
  \begin{tabular}{lrr}
    \toprule
    $(n_1, n_2, n_3)$  & $\real$-\textit{SCM} & $\cplx$-\textit{SCM} \\
    \midrule
    (\texttt{raw}) & $20,  50,  500$ & $10, 25, 250$ \\
    (\texttt{fft}) & $40, 100, 1000$ & $20, 50, 500$ \\
  %   \bottomrule
  % \end{tabular}
  % \begin{tabular}{lrr}
    \toprule
    $(n_1)$  & $\real$-\textit{TLDM} & $\cplx$-\textit{TLDM} \\
    \midrule
    (\texttt{raw}) & $4096$ & $2048$ \\
    (\texttt{fft}) & $8192$ & $4096$ \\
    \bottomrule
  \end{tabular}
\end{table}

@@

We use training from sec.~\ref{sub:fitting_and_sparsification} with epochs lasting for $12$,
$32$ and $50$ epochs, respectively. Early stopping is used only during the ``fine-tune''
stage. Preliminary experiments with the model without sparsification have shown that early
stopping almost always terminated the training process within the first $10-20$ epochs,
and that the validation performance peaked at about $10-15$ epochs, much sooner than $200$
epoch, used by \citet{trabelsi_deep_2017}. Besides stagewise training and shorter ``dense''
stage, other important deviations from their setup include:
\begin{itemize}
  \item shifting lower frequencies to the centre of the window to maintain locality after
  applying FFT (much like in sec.~\ref{sub:mnist_like_datasets})
  \item clipping $\ell_2$ norm of parameters' gradients to $0.05$
  \item faster learning rate reduction schedule (sec.~\ref{sub:fitting_and_sparsification})
\end{itemize}
Each full three-stage experiment is replicated five times to take random effects into account,
e.g. random network initialization prior to ``dense'' stage, randomization of batches within
each epoch during SGD, and stochasticity of Bayesian forward pass at ``sparsify'' stage.
